### 2021-07-14 src/paper/report.tex
Minor polishing of report.

At the core of our agent-based model \citep[][we review more literature in Supplementary
    Material~\ref{sec:literature_review}]{Aleta2020,Hinch2020} are physical contacts
between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between an infectious individual and somebody susceptible to
the disease bears the risk of transmitting the virus. Contacts occur in up to four networks:
Within the household, at work, at school, or in other settings (leisure activities,
grocery shopping, medical appointments, etc.). Some contacts recur regularly, others
occur at random. Empirical applications can take the population and household structure
from census data and the network-specific frequencies of
contacts from diary data measuring contacts before the pandemic
\citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies depend
on age and geographical location (see Supplementary
Material~\ref{subsec:assortativity}).

@@

    \floatfoot{\noindent \textit{Note:}
        A~description of the model can be found in Supplementary
        Material~\ref{sec:model}. Figure~\ref{fig:model_contacts_infections} shows the
        influence on an agent's contacts to other agents. Demographic characteristics
        set the baseline number of contacts in different networks ($\eta$). She may
        reduce the number of contacts due to NPIs, showing symptoms, or testing
        positively for SARS-CoV-2 ($\tau$). Infections may occur when a susceptible
        agent meets an infectious agent; the probability depends on the type of contact
        ($\beta_c$), on seasonality ($\kappa_{c}$), and on NPIs ($\rho_{c,\:t}$). If
        infected, the infection progresses as depicted in
        Figure~\ref{fig:model_disease_progression}. If rapid tests are available,
        agents' demand is modeled as in Figure~\ref{fig:model_rapid_tests}. All reasons
        trigger a test only for a fraction of individuals depending on an individual
        compliance parameter; the thresholds for triggering test demand differ across
        reasons and they may depend on calendar time ($\pi_{c,\:t}$ and $\tau_{c,\:t}$).
        Figure~\ref{fig:model_official_cases} shows the model of translating all
        infections in the simulated data to age-specific recorded infections. The model
        uses data on the aggregate share of recorded cases ($\psi$), on the share of
        positive PCR tests triggered by symptoms ($\chi_{symptom}$), and on the false
        positive rate of rapid tests ($p_{positive|infected,\;i,\:t}$). The lower part
        of the graph is relevant only for periods where rapid tests are available.}
\end{figure}

@@

We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
individual is infected or not; there is no uncertainty to the result. PCR tests require
at least one day to be processed and there are aggregate capacity constraints. In
contrast, rapid antigen tests
yield immediate results. Specificity and
sensitivity of these tests is set
according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
timing of the test relative to the onset of infectiousness. After a phase-in period, all
tests that are demanded will be performed. Figure~\ref{fig:model_rapid_tests} shows our
model for rapid test demand. Schools may require staff and students to be tested
regularly. Rapid tests may be offered by
employers to on-site workers. Individuals may demand tests for private reasons, which
include having plans to meet other people, showing symptoms of CoViD-19, and a household
member having tested positively for the virus. We endow each agent with an individual
compliance parameter. This parameter determines whether she takes up rapid
tests.\footnote{Positive test results or symptoms leads most individuals to reduce their
    contacts; this is why tests impact the actual contacts in
    Figure~\ref{fig:model_description}.}

Modelling a population of agents according to actual demographic characteristics means
that we can use a wide array of data to identify and estimate the model's many
parameters.\footnote{See Supplementary Material~\ref{sec:data_and_parameters} for a
    complete description.} Contact diaries yield pre-pandemic distributions of contacts for
different contact types and their assortativity by age group. Mobility data is used to
model the evolution of work contacts. School and daycare policies can be incorporated
directly from official directives. Administrative records on the number of tests,
vaccinations by age and region, and the prevalence of virus strains are generally
available. Surveys may ask about test offers, propensities to take them up, and past
tests. Other studies' estimates of the seasonality of infections can be incorporated
directly. The remaining parameters---most notably, these include infection probabilities
by contact network and the effects of some NPIs, see Supplementary
Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
matches features of the data \citep[see][for the general method]{McFadden1989}. In our
application, we keep the number of free parameters low in order to avoid overfitting.
The data features to be matched include official case numbers for each age group and
region, deaths, and the share of the B.1.1.7 strain.

The main issue with official case numbers is that they will contain only a fraction of
all infections. In the German case, this specifically amounts to positive PCR tests. We
thus model recorded cases as depicted in Figure~\ref{fig:model_official_cases}. We take
mortality-based\comment[id=HM]{added ``mortality-based'' -- correct?} aggregate
estimates of the share of detected cases and use data on the
share of PCR tests administered because of CoViD-19
symptoms. As the share of asymptomatic individuals
varies by age group, this gives us age-specific shares (see
Figure~\ref{fig:share_known_cases_by_age_group} for the share of detected cases by age
group over time in our model). Our estimates suggest that---in the absence of rapid
testing---the detection rate is 80\% higher on average for individuals above age 80
compared to school age children. Once rapid test become available, confirmation of a
positive result
is another reason leading to positive PCR tests.

@@

    \floatfoot{\noindent \textit{Note:} The blue line in
        Figure~\ref{fig:2021_scenarios_recorded} is the same as in
        Figure~\ref{fig:aggregated_fit} and refers to our baseline scenario, so does the
        blue line in Figure~\ref{fig:2021_scenarios_newly_infected}. The red lines refer
        to a situation where NPIs evolve as in the baseline scenario and the B.1.1.7
        variant is introduced in the same way; vaccinations, rapid tests, and
        seasonality remain at their January levels. The other scenarios turn each of
        these three factors on individually. The decompositions in
        Figures~\ref{fig:2021_scenarios_decomposition} and
        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values, which
        are explained more thoroughly in Appendix~\ref{sub:shapley_value}}
\end{figure}

Until mid-March, there is no visible difference between the different scenarios.
Seasonality hardly changes, and only few vaccinations and rapid tests were administered.
Even thereafter, the effect of the vaccination campaign is surprisingly small at first
sight. Whether considering recorded or total infections with only one channel active,
the final level is always the highest in case of the vaccination campaign (orange
lines). The Shapley value decomposition shows that vaccinations contribute 16\% to the
cumulative difference between scenarios. Reasons for the low share are the slow
start---it took until March~24th until 10\% of the population had received their first
vaccination, the 20\% mark was reached on April 19th---and the focus on older
individuals. These groups contribute less to the spread of the disease than others due
to a lower number of contacts. By the end of our study period, when first-dose
vaccination rates reached 43\% of the population, the numbers of new cases would have
started to decline. It is important to note that the initial focus of the campaign was
to prevent deaths and severe disease. Indeed, the case fatality was rate considerably
lower during the third wave when compared to the second (4.4\% between October and
February and 1.4\% between March and the end of May).\comment[id=K]{This needs a
source.}\comment[id=HM]{Didn't we calculate that ourselves at some point?}

Seasonality has a large effect in slowing the spread of SARS-CoV-2. By May 31, both
observed and recorded cases would be reduced by a factor of four if only seasonality
mattered. However, in this period, cases would have kept on rising throughout, just at a
much lower pace \citep[this is in line with results in][, which our seasonality measure
    is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be a
quantitatively important factor determining the evolution of the pandemic, explaining
most of the early changes and 43\% of the cumulative difference by the end of May.

@@

So why is rapid testing so effective? In order to shed more light on this question,
Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
scenario without rapid tests only and the main specification into the three channels for
rapid tests. Tests at schools have the smallest effect, which is largely explained by
schools not operating at full capacity during our period of study and the relatively
small number of students.\footnote{18\% of our population are in the education sector
    (pupils, teachers, etc.); 46\% are workers outside the education sector.} Almost 40\%
come from tests at the workplace. Despite the fact that rapid tests for private reasons
are phased in only in mid-March, they make up for more than half of the total effect.
The reason lies in the fact that a substantial share of these tests is driven by an
elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
up on a positive test of a household member. The latter is essentially a form of contact
tracing, which has been shown to be very effective \citep{Contreras2021,
    Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in Supplementary
Material~\ref{subsec:appendix_scenarios} shows that the same amount of rapid tests
administered randomly in the population would not nearly have been as effective.

Two of the most contentious NPIs concern schools and mandates to work from
home.\comment[id=K]{We should decide if we want to keep this section or prefer to show
some rapid test statistics.} In many countries, schools switched to remote instruction
during the first wave, so did Germany. After the summer break, they were operating at
full capacity with increased hygiene measures, before being closed again from
mid-December onward. Some states started opening them gradually in late February, but
operation at normal capacity did not resume until the beginning of June.
Figure~\ref{fig:school_scenarios} shows the effects of different policies regarding
schools starting at Easter, at which point rapid tests had become widely available. We
estimate the realized scenario to have essentially the same effect as a situation with
closed schools. Under fully opened schools with mandatory tests, total infections would
have been 6\% higher; this number rises to 20\% without tests. These effect sizes are
broadly in line with empirical studies \citep[e.g.,][]{Vlachos2021}.\comment[id=HM]{Need
a back-of the envelope calculation} To use another metric, the effective weekly
reproduction number differs by $0.018$ and $0.052$, respectively. In light of the large
negative effects school closures have on children and parents \citep{Luijten2021,
Melegari2021}---and in particular on those with low socio-economic status---these results
in conjunction with hindsight bias suggest that opening schools combined with a testing
strategy would have been beneficial. In other situations, and particular when rapid test
are not available at scale, trade-offs may well be different.

@@

    \floatfoot{\noindent \textit{Note:} Blue lines in both figures refer to our baseline
        scenario; they are the same as in
        Figure~\ref{fig:2021_scenarios_newly_infected}. Interventions start at Easter
        because there were no capacity constraints for rapid tests afterwards.}

@@

Our analysis has shown that during the transition to high levels of vaccination and
possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
estimated to have cost around 20~Euros per capita \citep{Wollmershauser2021}; retail
prices for rapid tests were below one Euro in early June 2021. While we do not
distinguish between self-administered rapid tests and point of care rapid tests, the
former are likely to play a larger role for indication-driven testing. Widespread
availability at low prices seems important. However, they rely on purely voluntary
participation in a non-public setting. The benefit of point-of-care rapid tests as a
precondition to participate in leisure activities as well as mandatory tests at the
workplace or at school come from screening the entire population. This is important
because disadvantaged groups are less likely to be reached by testing campaigns relying
on voluntary participation \citep[e.g.][]{StillmanTonin2021}; at the same time, these
groups have a higher risk to contract CoViD-19 \citep{KochInstitut2021a}. Mandatory
tests at school and at the workplace will extend more into these groups. The same goes
for individuals who exhibit a low level of compliance with CoViD-19-related regulations.
Compared to vaccinations, rapid testing programmes allow a much quicker roll-out, making
it arguably the most effective tool to contain the pandemic in the short run.

>>>>

At the core of our agent-based model \citep[][we review more literature in Supplementary
    Material~\ref{sec:literature_review}]{Aleta2020,Hinch2020} are physical contacts
between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between an infectious individual and somebody susceptible to
the disease bears the risk of transmitting the virus. Contacts occur in up to four networks:
 \comment[id=J]{I would replace networks by areas because we have more than four networks}
Within the household, at work, at school, or in other settings (leisure activities,
grocery shopping, medical appointments, etc.). Some contacts recur regularly, others
occur at random. Empirical applications can take the population and household structure
from census data and the network-specific frequencies of
contacts from diary data measuring contacts before the pandemic
\citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies depend
on age and geographical location (see Supplementary
Material~\ref{subsec:assortativity}).

@@

    \floatfoot{\noindent \textit{Note:}
        A~description of the model can be found in Supplementary
        Material~\ref{sec:model}. Figure~\ref{fig:model_contacts_infections} shows the
        influence of an agent's contacts to other agents on infections. Demographic characteristics
        set the baseline number of contacts in different networks ($\eta$). The agents may
        reduce the number of contacts due to NPIs, showing symptoms, or testing
        positively for SARS-CoV-2 ($\tau$). Infections may occur when a susceptible
        agent meets an infectious agent; the probability depends on the type of contact
        ($\beta_c$), on seasonality ($\kappa_{c}$), and on NPIs ($\rho_{c,\:t}$). If
        infected, the infection progresses as depicted in
        Figure~\ref{fig:model_disease_progression}. If rapid tests are available,
        agents' demand is modeled as in Figure~\ref{fig:model_rapid_tests}. All reasons
        trigger a test only for a fraction of individuals depending on an individual
        compliance parameter; the thresholds for triggering test demand differ across
        reasons and they may depend on calendar time ($\pi_{c,\:t}$ and $\tau_{c,\:t}$).
        Figure~\ref{fig:model_official_cases} shows the model of translating all
        infections in the simulated data to age-specific recorded infections. The model
        uses data on the aggregate share of recorded cases ($\psi$), the share of
        positive PCR tests triggered by symptoms ($\chi_{symptom}$), and the false
        positive rate of rapid tests ($p_{positive|infected,\;i,\:t}$). The lower part
        of the graph is relevant only for periods where rapid tests are available. All
        parameters are explained in Appendix\ref{sub:param_tables}.}
\end{figure}

@@

We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
individual is infected or not; there is no uncertainty to the result. PCR tests require
at least one day to be processed and there are aggregate capacity constraints. In
contrast, rapid antigen tests
yield immediate results. Specificity and
sensitivity of these tests is set
according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
timing of the test relative to the onset of infectiousness. After a phase-in period, all
tests that are demanded will be performed. Figure~\ref{fig:model_rapid_tests} shows our
model for rapid test demand. Schools may require staff and students to be tested
regularly. Rapid tests may be offered by
employers to on-site workers. Individuals may demand tests for private reasons, which
include having plans to meet other people, showing symptoms of CoViD-19, and a household
member having tested positively for the virus. We endow each agent with an individual
compliance parameter. This parameter determines whether she takes up rapid
tests.\footnote{Positive test results or symptoms lead most individuals to reduce their
    contacts; this is why tests impact the actual contacts in
    Figure~\ref{fig:model_description}.}

Modelling a population of agents according to actual demographic characteristics means
that we can use a wide array of data to identify and calibrate the model's many
parameters.\footnote{See Supplementary Material~\ref{sec:data_and_parameters} for a
    complete description.} Contact diaries yield pre-pandemic distributions of contacts for
different contact types and their assortativity by age group. Mobility data is used to
model the evolution of work contacts. School and daycare policies can be incorporated
directly from official directives. Administrative records on the number of tests,
vaccinations by age and region, and the prevalence of virus strains are generally
available. Surveys may ask about test offers, propensities to take them up, and past
tests. Other studies' estimates of the seasonality of infections can be incorporated
directly. The remaining parameters---most notably, these include infection probabilities
by contact network and the effects of some NPIs, see Supplementary
Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
matches features of the data \citep[see][for the general method]{McFadden1989}. In our
application, we keep the number of free parameters low in order to avoid overfitting.
The data features to be matched include official case numbers for each age group and
region, deaths, and the share of the B.1.1.7 strain.

The main issue with official case numbers is that they will contain only a fraction of
all infections. In the German case, this specifically amounts to positive PCR tests. We
thus model recorded cases as depicted in Figure~\ref{fig:model_official_cases}. We take
mortality-based aggregate estimates of the share of detected cases and use data on the
share of PCR tests administered because of CoViD-19
symptoms. As the share of asymptomatic individuals
varies by age group, this gives us age-specific shares (see
Figure~\ref{fig:share_known_cases_by_age_group}). Our estimates suggest that---in the
absence of rapid testing---the detection rate is 80\% higher on average for individuals
above age 80 compared to school age children. Once rapid test become available,
confirmation of a positive result is another reason leading to positive PCR tests.

@@

    \floatfoot{\noindent \textit{Note:} The blue line in
        Figure~\ref{fig:2021_scenarios_recorded} is the same as in
        Figure~\ref{fig:aggregated_fit} and refers to our baseline scenario, so does the
        blue line in Figure~\ref{fig:2021_scenarios_newly_infected}. The red lines refer
        to a situation where NPIs evolve as in the baseline scenario and the B.1.1.7
        variant is introduced in the same way; vaccinations, rapid tests, and
        seasonality remain at their January levels. The other scenarios turn each of
        these three factors on individually. The decompositions in
        Figures~\ref{fig:2021_scenarios_decomposition} and
        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values, which
        are explained more thoroughly in Appendix~\ref{sub:shapley_value}.
        For legibility reasons, all lines are rolling 7-day averages.}
\end{figure}

Until mid-March, there is no visible difference between the different scenarios.
Seasonality hardly changes, and only few vaccinations and rapid tests were administered.
Even thereafter, the effect of the vaccination campaign is surprisingly small at first
sight. Whether considering recorded or total infections with only one channel active,
the final level is always the highest in case of the vaccination campaign (orange
lines). The Shapley value decomposition shows that vaccinations contribute 16\% to the
cumulative difference between scenarios. Reasons for the low share are the slow
start---it took until March~24th until 10\% of the population had received their first
vaccination, the 20\% mark was reached on April 19th---and the focus on older
individuals. These groups contribute less to the spread of the disease than others due
to a lower number of contacts. By the end of our study period, when first-dose
vaccination rates reached 43\% of the population, the numbers of new cases would have
started to decline. It is important to note that the initial focus of the campaign was
to prevent deaths and severe disease. Indeed, the case fatality rate was considerably
lower during the third wave when compared to the second (4.4\% between October and
February and 1.4\% between March and the end of May).

Seasonality has a large effect in slowing the spread of SARS-CoV-2. By May 31, both
observed and total cases would be reduced by a factor of four if only seasonality
mattered. However, in this period, cases would have kept on rising throughout, just at a
much lower pace \citep[this is in line with results in][, which our seasonality measure
    is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be a
quantitatively important factor determining the evolution of the pandemic, explaining
most of the early changes and 43\% of the cumulative difference by the end of May.

@@

So why is rapid testing so effective? In order to shed more light on this question,
Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
scenario without rapid tests and the main specification into the three channels for
rapid tests. Tests at schools have the smallest effect, which is largely explained by
schools not operating at full capacity during our period of study and the relatively
small number of students.\footnote{18\% of our population are in the education sector
    (pupils, teachers, etc.); 46\% are workers outside the education sector.} Almost 40\%
come from tests at the workplace. Despite the fact that rapid tests for private reasons
are phased in only in mid-March, they make up for more than half of the total effect.
The reason lies in the fact that a substantial share of these tests is driven by an
elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
up on a positive test of a household member. The latter is essentially a form of contact
tracing, which has been shown to be very effective \citep{Contreras2021,
    Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in Supplementary
Material~\ref{subsec:appendix_scenarios} shows that the same amount of rapid tests
administered randomly in the population would not have been nearly as effective.

Two of the most contentious NPIs concern schools and mandates to work from
home. In many countries, schools switched to remote instruction
during the first wave, so did Germany. After the summer break, they were operating at
full capacity with increased hygiene measures, before being closed again from
mid-December onward. Some states started opening them gradually in late February, but
operation at normal capacity did not resume until the beginning of June.
Figure~\ref{fig:school_scenarios} shows the effects of different policies regarding
schools starting after Easter, at which point rapid tests had become widely available. We
estimate the realized scenario to have essentially the same effect as a situation with
closed schools. Under fully opened schools with mandatory tests, total infections would
have been 6\% higher; this number rises to 20\% without tests. These effect sizes are
broadly in line with empirical studies \citep[e.g.,][]{Vlachos2021}.\comment[id=HM]{Need
a back-of the envelope calculation} To use another metric, the effective weekly
reproduction number differs by $0.018$ and $0.052$, respectively. In light of the large
negative effects school closures have on children and parents \citep{Luijten2021,
Melegari2021}---and in particular on those with low socio-economic status---these results
in conjunction with hindsight bias suggest that opening schools combined with a testing
strategy would have been beneficial. In other situations, and particular when rapid test
are not available at scale, trade-offs may well be different.

@@

    \floatfoot{\noindent \textit{Note:} Blue lines in both figures refer to our baseline
        scenario; they are the same as in
        Figure~\ref{fig:2021_scenarios_newly_infected}. Interventions start at Easter
        because there were no capacity constraints for rapid tests afterwards.
        For legibility reasons, all lines are rolling 7-day averages.}

@@

Our analysis has shown that during the transition to high levels of vaccination and
possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
estimated to have cost around 20~Euros per capita \citep{Wollmershauser2021}; retail
 \comment[id=J]{Check this number and add to which time period it refers}
prices for rapid tests were below one Euro in early June 2021. While we do not
distinguish between self-administered rapid tests and point of care rapid tests, the
former are likely to play a larger role for indication-driven testing. Widespread
availability at low prices seems important. However, they rely on purely voluntary
participation in a non-public setting. The benefit of point-of-care rapid tests as a
precondition to participate in leisure activities as well as mandatory tests at the
workplace or at school come from screening the entire population. This is important
because disadvantaged groups are less likely to be reached by testing campaigns relying
on voluntary participation \citep[e.g.][]{StillmanTonin2021}; at the same time, these
groups have a higher risk to contract CoViD-19 \citep{KochInstitut2021a}. Mandatory
tests at school and at the workplace will extend more into these groups. The same goes
for individuals who exhibit a low level of compliance with CoViD-19-related regulations.
Compared to vaccinations, rapid testing programmes allow a much quicker roll-out, making
it arguably the most effective tool to contain the pandemic in the short run.
